---
title: "Feature Selection Workflow Report"
date: "`r Sys.Date()`"
params:
   rmd: "report.Rmd"
output:
  html_document
---

```{r setup, include=FALSE}
library(knitr)
library(htmltools)
knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE)
```
***

#

<h1 style="background-color:lightgray; color: black; text-align: center; padding-top:30px; padding-bottom: 30px; padding-left: 20px;">Summary</h1>

***

Values of the response variable are explained best with the important features selected using this workflow.\

Please ensure you understand the workflow and that the accuracy of the model is adequate for your purposes.
When you are using all your data without prior
hypotheses, you are performing *exploratory analysis* and should make this clear
when discussing any results.\

Additional figures and files can be found in the output directory.

***

```{r comment=""}
response_var = snakemake@config[["response_var"]]
drop_cols = snakemake@config[['drop_cols']]
constrain_rows = snakemake@config[['constrain_rows']]
random_effect = snakemake@config[['random_effect']]
drop_rows = snakemake@config[['drop_rows']]
iterations = snakemake@config[["iterations"]]
n_estimators = snakemake@config[["n_estimators"]]
output_directory = snakemake@config[["out"]]

cat(paste0("RESPONSE VARIABLE: ", response_var))
cat(paste0("RANDOM EFFECT: ", random_effect))
cat(paste0("FEATURES DROPPED: ", toString(drop_cols)))
cat(paste0("VALUES DROPED ({feature: value}): ", toString(drop_rows)))
cat(paste0("CONSTRAINED TO ROWS: ", toString(constrain_rows)))
cat(paste0("OUTPUT DIRECTORY: ", toString(output_directory)))

cat(paste0("This pipeline was used to identify important features predictive of ", response_var, " values.\nA random effect of ", random_effect, " was used to adjust for non-independence. There were ", n_estimators, " trees per Random Forest model. Mixed-effects software MERF was used with ", iterations, " iterations. (Include log file here)"))


```

<br>
<br>
<br>
<br>
<br>
</br>

#

***

<h1 style="background-color:#FFFACD; color: black; text-align: center; padding-top:30px; padding-bottom: 30px; padding-left: 20px;">Original Dataset</h1>

***

## Results {.tabset}

### Log and SHAP Summary Results

```{r out.width = 1100, out.height = 600, comment=""}
cat(readLines(snakemake@input[['original_log']]), sep = '\n')
include_graphics(snakemake@input[['original_shap']])
```

### BorutaSHAP Important Features

```{r out.width = 1100, out.height = 600, comment=""}
include_graphics(snakemake@input[['original_boruta']])
```

### Results PDF

```{r out.width = 1100, out.height = 600, comment=""}
include_graphics(snakemake@input[['original']])
```
<br>
<br>
<br>
<br>
<br>
</br>

#
***

<h1 style="background-color:cornflowerblue; color: black; text-align: center; padding-top:30px; padding-bottom: 30px; padding-left: 20px;">First Delta Dataset</h1>

***

## Results {.tabset}

### Log and SHAP Summary Results

```{r out.width = 1100, out.height = 600, comment=""}
cat(readLines(snakemake@input[['first_log']]), sep = '\n')
include_graphics(snakemake@input[['first_shap']])
```

### BorutaSHAP Important Features

```{r out.width = 1100, out.height = 600, comment=""}
include_graphics(snakemake@input[['first_boruta']])
```

### Results PDF

```{r out.width = 1100, out.height = 600, comment=""}
include_graphics(snakemake@input[['first']])
```
<br>
<br>
<br>
<br>
<br>
</br>

#

***

<h1 style="background-color:lightgreen; color: black; text-align: center; padding-top:30px; padding-bottom: 30px; padding-left: 20px;">Previous Delta Dataset</h1>

***

## Results {.tabset}

### Log and SHAP Summary Results

```{r out.width = 1100, out.height = 600, comment=""}
cat(readLines(snakemake@input[['previous_log']]), sep = '\n')
include_graphics(snakemake@input[['previous_shap']])
```

### BorutaSHAP Important Features

```{r out.width = 1100, out.height = 600, comment=""}
include_graphics(snakemake@input[['previous_boruta']])
```

### Results PDF

```{r out.width = 1100, out.height = 600, comment=""}
include_graphics(snakemake@input[['previous']])
```
<br>
<br>
<br>
<br>
<br>
</br>

#

***

<h1 style="background-color:#AB82FF; color: black; text-align: center; padding-top:30px; padding-bottom: 30px; padding-left: 20px;">Pairwise Delta Dataset</h1>
***


## Results {.tabset}

### Log and SHAP Summary Results

```{r out.width = 1100, out.height = 600, comment=""}
cat(readLines(snakemake@input[['pairwise_log']]), sep = '\n')
include_graphics(snakemake@input[['pairwise_shap']])
```

### BorutaSHAP Important Features

```{r out.width = 1100, out.height = 600, comment=""}
include_graphics(snakemake@input[['pairwise_boruta']])
```

### Results PDF

```{r out.width = 1100, out.height = 600, comment=""}
include_graphics(snakemake@input[['pairwise']])
```
<br>
<br>
<br>
<br>
<br>
</br>

#

***

#

<h1 style="background-color:lightgray; color: black; text-align: center; padding-top:30px; padding-bottom: 30px; padding-left: 20px;">Post-hoc tests</h1>

# Results {.tabset}

***

## Description

Linear models or linear mixed effects models were used for post hoc testing to determine if a linear relationship exists between selected feature and response variable.
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

## Results

```{r out.width = 1100, out.height = 600, comment=""}
htmltools::includeHTML(snakemake@input[['post_hoc']])
```

***

See the [Github repository](https://github.com/JTFouquier/snakemake-MERF "Still a work in progress!! :) ") for more information.

***

```{r}
# read in a table example
# df = read.csv(snakemake@input[['table_boruta']], sep = "\t")
# knitr::kable(df$important.features, format = "html")
```

<a download="report.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime='text/rmd', encoding='base64')`">Click here for R Markdown source file</a>

