---
title: "Feature Selection Report"
date: "Date: `r Sys.Date()`"
output:
  html_document
---

```{r setup, include=FALSE}
source("scripts/install.R")
source("scripts/colors.R")
package_list <- c("knitr", "kableExtra", "htmltools", "tidyverse",
"Hmisc", "jsonlite")
install_r_packages(package_list = package_list)
library(knitr)
library(htmltools)
library(tidyverse)
library(Hmisc)
library(jsonlite)
library(kableExtra)

knitr::opts_chunk$set(echo = FALSE, eval = TRUE, error = FALSE,
                      warning = FALSE, message = FALSE,
                      comment = "")
options(warn = -1)
```

```{r comment="", echo = FALSE}

response_var <- snakemake@config[["response_var"]]
random_effect <- snakemake@config[["random_effect"]]

df_mod <- snakemake@config[["df_mod"]]

iterations <- snakemake@config[["iterations"]]
n_estimators <- snakemake@config[["n_estimators"]]
max_features <- snakemake@config[["max_features"]]
version <- snakemake@config[["version"]]

borutashap_trials <- snakemake@config[["borutashap_trials"]]
borutashap_threshold <- snakemake@config[["borutashap_threshold"]]
borutashap_p <- snakemake@config[["borutashap_p"]]

enc_percent_threshold <- snakemake@config[["enc_percent_threshold"]]

original_path <- snakemake@config[["path_original"]]
first_path <- snakemake@config[["path_first"]]
previous_path <- snakemake@config[["path_previous"]]
pairwise_path <- snakemake@config[["path_pairwise"]]


pretty_table <- function(dataset_type, display_file) {
  setwd(snakemake@config[["out"]])
  df_path <- paste0(snakemake@config[[paste0("path_", dataset_type)]],
  paste0(dataset_type, display_file))
  if (!file.exists(df_path)) {
    cat("Analysis not completed")
  } else {
    df <- read.csv(df_path, sep = "\t", check.names = FALSE)
    if ("Failed Analysis" %in% colnames(df)) {
      cat("")
    } else {
      kable(df, format = "html", escape = FALSE) %>%
        kable_styling(
          font_size = 11
        )
    }
  }
}

url_tables <- function(df_path) {
  setwd(snakemake@config[["out"]])
  df <- read.csv(df_path, sep = "\t", check.names = FALSE)

  if ("Failed Analysis" %in% colnames(df)) {
    cat("Analysis not completed")
  } else {
  df$url <- lapply(df$url, function(url) {
    as.character(a(href = url, target = "_blank", url))
  })

  df_table <- df %>%
    mutate(url = cell_spec(url, "html", escape = FALSE))

  kable(df_table, format = "html", escape = FALSE) %>%
    kable_styling()
  }
}
```

```{r include = FALSE }

plot_pattern_display <- function(path_name, pattern, grep_arg) {
  setwd(snakemake@config[["out"]])
  files <- list.files(path = path_name, pattern = pattern)
  if (length(files) < 1) {
    return(FALSE)
  } else {
    my_plots <- grep(grep_arg, files, value = TRUE)
    for (f in my_plots) {
        cat("\n\n")
        cat(paste0("![](", path_name, f, ")"), "\n")
    }
  }
  return(TRUE)
}

# create the plot to show which datasets important features
# were occurring in (done now after building all models)

# after variables are encoded, remove very low occurring ones if wanted.
if (enc_percent_threshold == "0") {

  enc_percent_threshold_print <- paste0("low occuring categorical",
  " values were not removed")

} else {

  enc_percent_threshold_print <- paste0("low occurring categorical values were removed if found in less than ", enc_percent_threshold, "% of samples") # nolint

}

# SHAP figure text


shap_description <- function(fig_flag){
  if (fig_flag == FALSE) {
    # no SHAP plots, no need for legend
    rmd_text <- c("")
  } else if (fig_flag == TRUE) {
    rmd_text <- c(
      "**SHAP summary beeswarm plots of feature influence on the machine learning prediction of response values.** Each point represents one sample, and the horizontal position indicates impact on the response as indicated on the x-axis. Points to the left indicate a negative impact, and points to the right indicate a positive impact. The colors represent feature values, where red is larger and blue is smaller. For binary encoded features ('ENC') red is yes[1] and blue is no[0]. SHAP is generally an improvement upon other importance scores because it provides information about both rank (how helpful the feature was compared to other features [y-axis]) and impact (a positive or negative impact on response values [x-axis]). If multiple figures are shown, scales may vary with a maximum of ten features per plot."
      )
  }
  shap_description <- paste(rmd_text, collapse = "\n")
  cat(shap_description)
}
```

#### Analyst: `r toString(snakemake@config[["analyst"]])`

***

<h1 style="background-color:`r color_summary`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">EXPLANA Description and Analysis Summary</h1>

***

#### EXPLANA Description

EXPLANA assists with selecting which features, out of many input features, that are best related to a response variable by employing well-known machine-learning algorithms combined with novel methods.

Input features and response variable can be numerical, categorical, or from non-normal data distributions. The relationships between selected features and the response can be linear, or complex, non-linear relationships. An interpretable report is generated that textually and graphically summarizes results for streamlined exploratory analysis and hypothesis generation.

For longitudinal datasets, changes in features for each study identifier, often subjects, are calculated using different reference points to obtain delta datasets (First, Previous and Pairwise delta datasets). This is important because features in longitudinal studies can carry varying degrees of importance between models built using different reference points. This workflow automates feature selection from several models built from these datasets.

Please ensure you understand the workflow and that the percent variation explained (using out-of-bag (OOB) scores) is adequate for your purposes.

When you are using data without prior hypotheses, you are performing *exploratory analysis* and should make this clear.

***

## Response Variable: *`r response_var`*

#### Analysis Notes: `r toString(snakemake@config[["analysis_notes"]])`

***

### {.tabset}

#### Analysis Summary

**Random effect:** ``r random_effect``

**Dataset Modifications:**

```{r results='asis'}
cat(df_mod)
```

***

##### Methods

Feel free to copy the following text, including citation information, for use in methods to ensure reproducibility:

EXPLANA was used for exploratory analysis to identify important features predictive of the response variable, **``r response_var``**.

Categorical variables were binary encoded and **``r enc_percent_threshold_print``**. A random effect of **``r random_effect``** was used to adjust for non-independence (repeated measurements) if needed. There were **``r n_estimators` trees`** used per Random Forest model with a max feature fraction of **``r max_features``**  of the input features for each split in a decision tree. If mixed effects Random Forests were needed,
**``r iterations` iterations`** were performed. BorutaSHAP was used with **``r borutashap_trials` trials`** to find features that perform repeatedly better than shuffled versions of all input features. Selected features were considered important if they performed better than **``r borutashap_threshold`%`** of the importance score of the best shuffled feature with a p-value of **``r borutashap_p`%`**.

These methods were copy/pasted from an EXPLANA (version: ``r version``) feature selection report created on ``r Sys.Date()`` (see github.com for code and additional documentation). Copyright 2023 Regents of the University of Colorado. All rights reserved.

***

#### Input Datasets

```{r comment=""}
ds_print <- snakemake@config[["input_datasets"]] %>%
  toJSON %>%
  prettify()
cat(ds_print)
```

### Directory Links

[⇨ Full Analysis Directory | ](.){target="_blank"}
[⇨ Original Dataset | ](./`r original_path`){target="_blank"}
[⇨ First Delta Dataset | ](./`r first_path`){target="_blank"}
[⇨ Previous Delta Dataset | ](./`r previous_path`){target="_blank"}
[⇨ Pairwise Delta Dataset](./`r pairwise_path`){target="_blank"}

***

### Results Summary {.tabset}

#### Selected Features by Dataset

```{r comment=""}
pretty_table("", "summary-log-table.txt")
```

***

```{r fig.align = "center"}
setwd(snakemake@config[["out"]])
include_graphics("important-feature-occurrences.svg")
```

**Selected features from models built with Original and, for longitudinal analyses, First, Previous and Pairwise delta datasets.** Average impact on response is displayed for positive instances of encoded variables (indicated with "ENC"). For numerical variables, the relationship to the response variable is more complex to determine, and indicated with "+/-". Grey boxes indicate features included in the model for each dataset. Long feature names may be truncated and indicated with ellipses. For comprehensive list of input features, see model details.

#### Interpretation/Literature Search

```{r}
url_tables("urls.txt")
```

#

***

<h1 style="background-color:`r color_original`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">Original Dataset</h1>

***

[⇨ Open File Directory](./`r original_path`){target="_blank"}

## {.tabset}

### Model Summary

```{r comment=""}
pretty_table("original", "-log-df.txt")
```

***

```{r results="asis", comment=""}
fig_flag = plot_pattern_display(path_name = original_path, pattern = ".svg",
                                grep_arg = "SHAP-summary-beeswarm")
```

```{r results='asis', echo=FALSE}
shap_description(fig_flag)
```

### Selected Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("original", "-boruta-important.txt")
```

### Feature Stats

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("original", "-feature-stats.txt")
```

### Input Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("original", "-input-features.txt")
```

### Figures

[⇨ Open PDF in new window](./`r original_path`original.pdf){target="_blank"}

```{r out.width = 1000, out.height = 600, comment=""}
setwd(snakemake@config[["out"]])
include_graphics(paste0(original_path, "original.pdf"))
```

### Interpretation/Literature Search

```{r}
url_tables(paste0(original_path, "original-urls.txt"))
```

### Log

```{r comment=""}
setwd(snakemake@config[["out"]])
cat(readLines(paste0(original_path, "original-log.txt")), sep = "\n")
```

##

***

<h1 style="background-color:`r color_first`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">First Delta Dataset</h1>

***

[⇨ Open File Directory](./`r first_path`){target="_blank"}

## Results {.tabset}

### Model Summary

```{r comment=""}
pretty_table("first", "-log-df.txt")
```

***

```{r results="asis", comment=""}
fig_flag = plot_pattern_display(path_name = first_path, pattern = ".svg",
                     grep_arg = "SHAP-summary-beeswarm")
```

```{r results='asis', echo=FALSE}
shap_description(fig_flag)
```

### Selected Features

```{r out.width = 1000, out.height = 600}
pretty_table("first", "-boruta-important.txt")
```

### Feature Stats

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("first", "-feature-stats.txt")
```

### Input Features

```{r out.width = 1000, out.height = 600}
pretty_table("first", "-input-features.txt")
```

### Figures

[⇨ Open PDF in new window](./`r first_path`first.pdf){target="_blank"}

```{r out.width = 1100, out.height = 600, comment=""}
setwd(snakemake@config[["out"]])
include_graphics(paste0(first_path, "first.pdf"))
```

### Interpretation/Literature Search

```{r}
url_tables(paste0(first_path, "first-urls.txt"))
```

### Log

```{r comment=""}
setwd(snakemake@config[["out"]])
cat(readLines(paste0(first_path, "first-log.txt")), sep = "\n")
```

##

***

<h1 style="background-color:`r color_previous`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">Previous Delta Dataset</h1>

***

[⇨ Open File Directory](./`r previous_path`){target="_blank"}

## {.tabset}

### Model Summary

```{r comment=""}
pretty_table("previous", "-log-df.txt")
```

***

```{r results="asis", comment=""}
fig_flag = plot_pattern_display(path_name = previous_path, pattern = ".svg",
                                grep_arg = "SHAP-summary-beeswarm")
```

```{r results='asis', echo=FALSE}
shap_description(fig_flag)
```

### Selected Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("previous", "-boruta-important.txt")
```

### Feature Stats

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("previous", "-feature-stats.txt")
```

### Input Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("previous", "-input-features.txt")
```

### Figures

[⇨ Open PDF in new window](./`r previous_path`previous.pdf){target="_blank"}

```{r out.width = 1000, out.height = 600, comment=""}
setwd(snakemake@config[["out"]])
include_graphics(paste0(previous_path, "previous.pdf"))
```

### Interpretation/Literature Search

```{r}
url_tables(paste0(previous_path, "previous-urls.txt"))
```

### Log

```{r comment=""}
setwd(snakemake@config[["out"]])
cat(readLines(paste0(previous_path, "previous-log.txt")), sep = "\n")
```

##

***

<h1 style="background-color:`r color_pairwise`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">Pairwise Delta Dataset</h1>

***

[⇨ Open File Directory](./`r pairwise_path`){target="_blank"}

## {.tabset}

### Model Summary

```{r comment=""}
pretty_table("pairwise", "-log-df.txt")
```

***

```{r results="asis", comment=""}
fig_flag = plot_pattern_display(path_name = pairwise_path, pattern = ".svg",
                                grep_arg = "SHAP-summary-beeswarm")
```

```{r results='asis', echo=FALSE}
shap_description(fig_flag)
```

### Selected Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("pairwise", "-boruta-important.txt")
```

### Feature Stats

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("pairwise", "-feature-stats.txt")
```

### Input Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("pairwise", "-input-features.txt")
```

### Figures

[⇨ Open PDF in new window](./`r pairwise_path`pairwise.pdf){target="_blank"}

```{r out.width = 1000, out.height = 600, comment=""}
setwd(snakemake@config[["out"]])
include_graphics(paste0(pairwise_path, "pairwise.pdf"))
```

### Interpretation/Literature Search

```{r}
url_tables(paste0(pairwise_path, "pairwise-urls.txt"))
```

### Log

```{r comment=""}
setwd(snakemake@config[["out"]])
cat(readLines(paste0(pairwise_path, "pairwise-log.txt")), sep = "\n")
```

##

***

See the [Github repository](https://github.com/JTFouquier/snakemake-MERF "Still a work in progress!! :) ") for more information.

***

```{r}
# read in a table example
# df = read.csv(snakemake@input[["table_boruta"]], sep = "\t")
# knitr::kable(df$important_features, format = "html")
```
