---
title: "Feature Selection Report"
date: "Date: `r Sys.Date()`"
output:
  html_document
---

```{r setup, include=FALSE}
source("scripts/install.R")
source("scripts/colors.R")
package_list <- c("knitr", "kableExtra", "htmltools", "tidyverse",
"Hmisc", "jsonlite")
install_r_packages(package_list = package_list)
library(knitr)
library(htmltools)
library(tidyverse)
library(Hmisc)
library(jsonlite)
library(kableExtra)

knitr::opts_chunk$set(echo = FALSE, eval = TRUE, error = FALSE,
                      warning = FALSE, message = FALSE,
                      comment = "")
options(warn = -1)
```

```{r comment="", echo = FALSE}

response_var <- snakemake@config[["response_var"]]
random_effect <- snakemake@config[["random_effect"]]

df_mod <- snakemake@config[["df_mod"]]

iterations <- snakemake@config[["iterations"]]
n_estimators <- snakemake@config[["n_estimators"]]
max_features <- snakemake@config[["max_features"]]
version <- snakemake@config[["version"]]

borutashap_trials <- snakemake@config[["borutashap_trials"]]
borutashap_threshold <- snakemake@config[["borutashap_threshold"]]
borutashap_p <- snakemake@config[["borutashap_p"]]

enc_percent_threshold <- snakemake@config[["enc_percent_threshold"]]

original_path <- snakemake@config[["path_original"]]
first_path <- snakemake@config[["path_first"]]
previous_path <- snakemake@config[["path_previous"]]
pairwise_path <- snakemake@config[["path_pairwise"]]


pretty_table <- function(dataset_type, display_file) {
  setwd(snakemake@config[["out"]])
  df_path <- paste0(snakemake@config[[paste0("path_", dataset_type)]],
  paste0(dataset_type, display_file))
  if (!file.exists(df_path)) {
    cat("Analysis not completed")
  } else {
    df <- read.csv(df_path, sep = "\t", check.names = FALSE)
    if ("Failed Analysis" %in% colnames(df)) {
      cat("")
    }
    else {
      kable(df, align = "c", escape = TRUE) %>%
        kable_styling(
          font_size = 11
        )
    }
  }
}

url_tables <- function(df_path) {
  setwd(snakemake@config[["out"]])
  df <- read.csv(df_path, sep = "\t", check.names = FALSE)

  if ("Failed Analysis" %in% colnames(df)) {
    cat("Analysis not completed")
  } else {
  df$url <- lapply(df$url, function(url) {
    as.character(a(href = url, target = "_blank", url))
  })

  df_table <- df %>%
    mutate(url = cell_spec(url, "html", escape = FALSE))

  kable(df_table, format = "html", escape = FALSE) %>%
    kable_styling(
      font_size = 10
    )
  }
}

url_statement = "The following links can help with hypothesis generation. Names of variables likely need modification."

```

```{r include = FALSE }

plot_pattern_display <- function(path_name, pattern, grep_arg) {
  setwd(snakemake@config[["out"]])
  files <- list.files(path = path_name, pattern = pattern)
  if (length(files) < 1) {
    return(FALSE)
  } else {
    my_plots <- grep(grep_arg, files, value = TRUE)
    for (f in my_plots) {
        cat("\n\n")
        cat(paste0("![](", path_name, f, ")"), "\n")
    }
  }
  return(TRUE)
}

# create the plot to show which datasets important features
# were occurring in (done now after building all models)

# after variables are encoded, remove very low occurring ones if wanted.
if (enc_percent_threshold == "0") {

  enc_percent_threshold_print <- paste0("low occuring categorical",
  " values were not removed")

} else {

  enc_percent_threshold_print <- paste0("low occurring categorical values were removed if found in less than ", enc_percent_threshold, "% of samples") # nolint

}

# SHAP figure text


shap_description <- function(fig_flag){
  if (fig_flag == FALSE) {
    # no SHAP plots, no need for legend
    rmd_text <- c("")
  } else if (fig_flag == TRUE) {
    rmd_text <- c(
      "**SHAP summary beeswarm plots of feature influence on the machine learning prediction of response values.** Each point represents one sample, and the horizontal position indicates impact on the response as indicated on the x-axis. Points to the left indicate a negative impact, and points to the right indicate a positive impact. The colors represent the selected feature values, where red is larger and blue is smaller. For binary encoded features ('ENC') red is yes[1] and blue is no[0]. SHAP is generally an improvement upon other importance scores because it provides information about both rank (how helpful the feature was compared to other features [y-axis]) and impact (a positive or negative impact on response values [x-axis]). If multiple figures are shown, scales may vary with a maximum of ten features per plot."
      )
  }
  shap_description <- paste(rmd_text, collapse = "\n")
  cat(shap_description)
}
```

#### Analyst: `r toString(snakemake@config[["analyst"]])`

***

<h1 style="background-color:`r color_summary`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">EXPLANA Description and Analysis Summary</h1>

***

#### EXPLANA Description

EXPLANA uses established machine-learning methods combined with innovative techniques to identify the most relevant features, from a variety of input features, that correlate with a response variable.

Input features and response variables can be numerical, categorical, or from non-normal data distributions. The relationships between selected features and the response can be linear or complex, non-linear relationships.

For longitudinal datasets, changes in features for each study identifier, often subjects, are calculated using different reference points to obtain delta datasets (*First*, *Previous* and *Pairwise* delta datasets). This is important because features in longitudinal studies can carry varying degrees of importance between models built using different reference points. EXPLANA automates feature selection from several models built using these datasets. This report was generated to textually and graphically summarize exploratory analysis and aid hypothesis generation.

Please ensure you understand the workflow, parameters/decisions, and that the percent variation explained (using out-of-bag (OOB) scores) is adequate for your purposes.

When you are using data without prior hypotheses, you are performing *exploratory analysis* and should make this clear when communicating results.

***

### Response Variable: *`r response_var`*

#### Analysis Notes: `r toString(snakemake@config[["analysis_notes"]])`

***

### {.tabset}

#### Analysis Summary

**Random effect:** ``r random_effect``

**Dataset Modifications:**

```{r results='asis'}
cat(df_mod)
```

***

##### Methods

Feel free to use the following text, including citation information, for use in methods to ensure reproducibility:

EXPLANA was used for exploratory analysis to identify important features related to the response variable, **``r response_var``**.

A random effect of **``r random_effect``** was used to adjust for non-independence (repeated measurements) if needed. There were **``r n_estimators``** trees used per Random Forest model with a max feature fraction of **``r max_features``**  of the input features for each split per decision tree in the forest. If mixed effects Random Forests were needed,
**``r iterations``** iterations were performed. BorutaSHAP was used to find features that perform repeatedly better than shuffled versions of all input features. Features were considered important if they performed better than **``r borutashap_threshold``****%** of the SHAP importance score of the best shuffled feature using **``r borutashap_trials``** trials, p=**``r borutashap_p``**. Categorical variables were binary encoded and **``r enc_percent_threshold_print``**.

These methods are from an EXPLANA feature selection report (version: **``r version``**) created on **``r Sys.Date()``**. See github.com for code and additional documentation. Copyright 2023 Regents of the University of Colorado. All rights reserved.

***

#### Input Datasets

```{r comment=""}
ds_print <- snakemake@config[["input_datasets"]] %>%
  toJSON %>%
  prettify()
cat(ds_print)
```

### Directory Links

[⇨ Full Analysis Directory | ](.){target="_blank"}
[⇨ *Original* Dataset | ](./`r original_path`){target="_blank"}
[⇨ *First* Delta Dataset | ](./`r first_path`){target="_blank"}
[⇨ *Previous* Delta Dataset | ](./`r previous_path`){target="_blank"}
[⇨ *Pairwise* Delta Dataset](./`r pairwise_path`){target="_blank"}

***

### Results Summary {.tabset}

#### Selected Features by Dataset

```{r comment=""}
pretty_table("", "summary-log-table.txt")
```

***

**Selected feature ranks from models built using *Original* and, for longitudinal analyses, *First*, *Previous* and *Pairwise* delta datasets.** Selected features are shown in black and labeled with feature rank. For true/positive instances of categorical variables (indicated with "ENC" after encoding), average impact on response is shown after the rank. For numerical features, impact is not shown because the feature relationship to response can be complex, requiring further post-hoc tests or inspection of SHAP dependence plots for additional insight. Empty grey boxes indicate features included in the model for a dataset, but not selected. Long feature names may be truncated and indicated with ellipses. A comprehensive list of input features can be found in model details below.

```{r fig.align = "center"}
setwd(snakemake@config[["out"]])
include_graphics("important-feature-occurrences.svg")
```

#### Interpretation/Literature Search

```{r comment=""}
cat(url_statement)
```

```{r}
url_tables("urls.txt")
```

#

***

<h1 style="background-color:`r color_original`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">Original Dataset</h1>

***

[⇨ Open File Directory](./`r original_path`){target="_blank"}

## {.tabset}

### Model Summary

```{r comment=""}
pretty_table("original", "-log-df.txt")
```

***

```{r results="asis", comment=""}
fig_flag = plot_pattern_display(path_name = original_path, pattern = ".svg",
                                grep_arg = "SHAP-summary-beeswarm")
```

```{r results='asis', echo=FALSE}
shap_description(fig_flag)
```

### Selected Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("original", "-boruta-important.txt")
```

### Feature Stats

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("original", "-feature-stats.txt")
```

### Input Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("original", "-input-features.txt")
```

### Figures

[⇨ Open PDF in new window](./`r original_path`original.pdf){target="_blank"}

```{r out.width = 1000, out.height = 600, comment=""}
setwd(snakemake@config[["out"]])
include_graphics(paste0(original_path, "original.pdf"))
```

### Interpretation/Literature Search

```{r comment=""}
cat(url_statement)
```

```{r}
url_tables(paste0(original_path, "original-urls.txt"))
```

### Log

```{r comment=""}
setwd(snakemake@config[["out"]])
cat(readLines(paste0(original_path, "original-log.txt")), sep = "\n")
```

##

***

<h1 style="background-color:`r color_first`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">First Delta Dataset</h1>

***

[⇨ Open File Directory](./`r first_path`){target="_blank"}

## Results {.tabset}

### Model Summary

```{r comment=""}
pretty_table("first", "-log-df.txt")
```

***

```{r results="asis", comment=""}
fig_flag = plot_pattern_display(path_name = first_path, pattern = ".svg",
                     grep_arg = "SHAP-summary-beeswarm")
```

```{r results='asis', echo=FALSE}
shap_description(fig_flag)
```

### Selected Features

```{r out.width = 1000, out.height = 600}
pretty_table("first", "-boruta-important.txt")
```

### Feature Stats

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("first", "-feature-stats.txt")
```

### Input Features

```{r out.width = 1000, out.height = 600}
pretty_table("first", "-input-features.txt")
```

### Figures

[⇨ Open PDF in new window](./`r first_path`first.pdf){target="_blank"}

```{r out.width = 1100, out.height = 600, comment=""}
setwd(snakemake@config[["out"]])
include_graphics(paste0(first_path, "first.pdf"))
```

### Interpretation/Literature Search

```{r comment=""}
cat(url_statement)
```

```{r}
url_tables(paste0(first_path, "first-urls.txt"))
```

### Log

```{r comment=""}
setwd(snakemake@config[["out"]])
cat(readLines(paste0(first_path, "first-log.txt")), sep = "\n")
```

##

***

<h1 style="background-color:`r color_previous`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">Previous Delta Dataset</h1>

***

[⇨ Open File Directory](./`r previous_path`){target="_blank"}

## {.tabset}

### Model Summary

```{r comment=""}
pretty_table("previous", "-log-df.txt")
```

***

```{r results="asis", comment=""}
fig_flag = plot_pattern_display(path_name = previous_path, pattern = ".svg",
                                grep_arg = "SHAP-summary-beeswarm")
```

```{r results='asis', echo=FALSE}
shap_description(fig_flag)
```

### Selected Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("previous", "-boruta-important.txt")
```

### Feature Stats

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("previous", "-feature-stats.txt")
```

### Input Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("previous", "-input-features.txt")
```

### Figures

[⇨ Open PDF in new window](./`r previous_path`previous.pdf){target="_blank"}

```{r out.width = 1000, out.height = 600, comment=""}
setwd(snakemake@config[["out"]])
include_graphics(paste0(previous_path, "previous.pdf"))
```

### Interpretation/Literature Search

```{r comment=""}
cat(url_statement)
```

```{r}
url_tables(paste0(previous_path, "previous-urls.txt"))
```

### Log

```{r comment=""}
setwd(snakemake@config[["out"]])
cat(readLines(paste0(previous_path, "previous-log.txt")), sep = "\n")
```

##

***

<h1 style="background-color:`r color_pairwise`; color: black; text-align: center; padding-top:20px; padding-bottom: 20px; padding-left: 20px;">Pairwise Delta Dataset</h1>

***

[⇨ Open File Directory](./`r pairwise_path`){target="_blank"}

## {.tabset}

### Model Summary

```{r comment=""}
pretty_table("pairwise", "-log-df.txt")
```

***

```{r results="asis", comment=""}
fig_flag = plot_pattern_display(path_name = pairwise_path, pattern = ".svg",
                                grep_arg = "SHAP-summary-beeswarm")
```

```{r results='asis', echo=FALSE}
shap_description(fig_flag)
```

### Selected Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("pairwise", "-boruta-important.txt")
```

### Feature Stats

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("pairwise", "-feature-stats.txt")
```

### Input Features

```{r out.width = 1000, out.height = 600, comment=""}
pretty_table("pairwise", "-input-features.txt")
```

### Figures

[⇨ Open PDF in new window](./`r pairwise_path`pairwise.pdf){target="_blank"}

```{r out.width = 1000, out.height = 600, comment=""}
setwd(snakemake@config[["out"]])
include_graphics(paste0(pairwise_path, "pairwise.pdf"))
```

### Interpretation/Literature Search

```{r comment=""}
cat(url_statement)
```

```{r}
url_tables(paste0(pairwise_path, "pairwise-urls.txt"))
```

### Log

```{r comment=""}
setwd(snakemake@config[["out"]])
cat(readLines(paste0(pairwise_path, "pairwise-log.txt")), sep = "\n")
```

##

***

See the [Github repository](https://github.com/JTFouquier/snakemake-MERF "Still a work in progress!! :) ") for more information.

***

```{r}
# read in a table example
# df = read.csv(snakemake@input[["table_boruta"]], sep = "\t")
# knitr::kable(df$important_features, format = "html")
```
