---
title: "Feature Selection Workflow Final Report"
date: "`r Sys.Date()`"
params:
   rmd: "report.Rmd"
output:
  html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE)
```
***

# Summary:

***

Values of the response variable are explained best with the important features selected using this workflow. Additional figures and files
can be found in the output directory. Please ensure you understand the workflow and that the accuracy of the model is adequate for your purposes.
When you are using all your data without prior
hypotheses, you are performing *exploratory analysis* and should make this clear
when discussing any results.

***

## Arguments:

***

```{r comment=""}
response_var = snakemake@config[["response_var"]]
drop_cols = snakemake@params[['drop_cols']]
constrain_rows = snakemake@params[['constrain_rows']]
random_effect = snakemake@config[['random_effect']]
drop_rows = snakemake@params[['drop_rows']]
iterations = snakemake@config[["iterations"]]

cat(paste0("RESPONSE VARIABLE: ", response_var))
cat(paste0("RANDOM EFFECT: ", random_effect))
cat(paste0("FEATURES DROPPED: ", toString(drop_cols)))
cat(paste0("VALUES DROPED ({feature: value}): ", toString(drop_rows)))
cat(paste0("CONSTRAINED TO ROWS: ", toString(constrain_rows)))


cat(paste0("This pipeline was used to identify important features predictive of ", response_var, " values. A random effect of ", random_effect, " was used to adjust for non-independence. Mixed-effects software MERF was used with ", iterations, " iterations. (Include log file here)"))


```

***

See the [Github repository](https://github.com/JTFouquier/snakemake-MERF "Still a work in progress!! :) ") for more information.

***

# Original Data

***

## Original Data: Log File

***


```{r comment=""}
cat(readLines(snakemake@input[['original_log']]), sep = '\n')
```

***

## Original Data: Overview of Results

***

```{r, out.width = "100%", out.height = 600}
include_graphics(snakemake@input[['original']])
```

***

## Original Data: SHAP summary

***

`r include_graphics(snakemake@input[['original_shap']])`

***

## Original Data: BorutaSHAP

***

`r include_graphics(snakemake@input[['original_boruta']])`




***

# First Delta Dataset


***

## First Delta Dataset: Log File

***

```{r comment=""}
cat(readLines(snakemake@input[['first_log']]), sep = '\n')
```
***

## First Delta Dataset: Overview of Results

***


```{r, out.width = "100%", out.height = 600}
include_graphics(snakemake@input[['first']])
```

***

## First Delta Dataset: SHAP summary

***

`r include_graphics(snakemake@input[['first_shap']])`

***

## First Delta Dataset: BorutaSHAP

***

`r include_graphics(snakemake@input[['first_boruta']])`




***

# Previous Delta Dataset

***

## Previous Delta Dataset: Log File

***

```{r comment=""}
cat(readLines(snakemake@input[['previous_log']]), sep = '\n')
```
## Previous Delta Dataset: Overview of Results

***


```{r, out.width = "100%", out.height = 600}
include_graphics(snakemake@input[['previous']])
```

***

## Previous Delta Dataset: SHAP summary

***

`r include_graphics(snakemake@input[['previous_shap']])`

***

## Previous Delta Dataset: BorutaSHAP

***

`r include_graphics(snakemake@input[['previous_boruta']])`



***

# Pairwise Delta Dataset

***

## Pairwise Delta Dataset: Log File

***

```{r comment=""}
cat(readLines(snakemake@input[['pairwise_log']]), sep = '\n')
```
***

## Pairwise Delta Dataset: Overview of Results

***

```{r, out.width = "100%", out.height = 600}
include_graphics(snakemake@input[['pairwise']])
```

***

## Pairwise Delta Dataset: SHAP summary

***

`r include_graphics(snakemake@input[['pairwise_shap']])`

***

## Pairwise Delta Dataset: BorutaSHAP

***

`r include_graphics(snakemake@input[['pairwise_boruta']])`

***











## Directed acyclic graph (DAG) showing workflow
`r include_graphics(snakemake@input[['dag_plot']])`

RF iters used TODO:
`r snakemake@params[['iters']]`

```{r}
# read in a table example
# df = read.csv(snakemake@input[['table_boruta']], sep = "\t")
# knitr::kable(df$important.features, format = "html")
```

<a download="report.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime='text/rmd', encoding='base64')`">Click here for R Markdown source file</a>

